---
title: "Midterm_Inclass"
author: "Emil Zetterquist"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own. I have not communicated with or worked with anyone else on this exam.”

Question 1:

Part a: data, stat, mapping coordinate function.

Part b: Realtive size is something that people notice pretty easily. By arranging the data by size it makes it easier for the viewer to draw coclusions abou the data.

Part: It¨s kinda the same thing. When using bar plots, size is one of the most important visualisations. By not including 0 in the plot, we can easily mislead the viewer into thinking that the relative size of the bars are bigger then they actually are.


<br>

Question 3: I've learned the importance of finding the right colour scale for your data in order for the viewer to properly see what you're trying to show. 


Questionn 4: 

part a:

```{r}
library(maps)
library(tidyverse)
library(here)

election_df <- read_csv(here("data/2020_county_pres.csv")) %>%
  group_by(state_name) %>%
  summarise(total_gop = sum(votes_gop),
            total_dem = sum(votes_dem)) %>%
  mutate(percent_gop = 100 * total_gop / (total_gop + total_dem)) %>%
  mutate(state_name = str_to_lower(state_name))


state_df <- ggplot2::map_data("state")

state_full <- left_join(state_df, election_df, by = c("region" = "state_name"))

ggplot(data = state_full, aes(x = long, y = lat, group = group)) +
  geom_polygon(colour = "black", aes(fill = percent_gop)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme_void() +
  scale_fill_viridis_b()
```

part b: beacuse the data needs a lot of tiny rows of coordniates in order to draw the map


Question 5: 

```{r}
library(shiny)
library(ggrepel)
alc_df <- read_csv(here("data/alcohol.csv")) 
var_choice <- names(alc_df)[2:4]


ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("countryselect",
                   label = "Choose a Country",
                   choices = levels(factor(alc_df$country))),
    radioButtons("varchoice",
                 label = "chose a stat",
                 choices = var_choice)),
  mainPanel(plotOutput(outputId = "scat"),
            plotOutput(outputId = "hist")))
)

server <- function(input, output, session) {
  
  alc_sub <- reactive({
    alc_df %>% filter(country == input$countryselect)
  })
  
  scat_plot <- reactive({
    ggplot(alc_df, aes(x = beer_servings, y = wine_servings)) +
  geom_point() +
  geom_label_repel(data = alc_sub, aes(label = country)) +
  geom_point(data = alc_sub, size = 3, shape = 1)
  })
  
  output$scat <- renderPlot({
    scat_plot()
  })
  
  hist_plot <- reactive({
    ggplot(alc_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = 19) +
      theme_grey(base_size = 22)
  })
  
  output$hist <- renderPlot({
    hist_plot()
  })
}

shinyApp(ui, server)

```

<br>

Question 7:

```{r}
wta_df <- read_csv(here("data/wta_matches_2019.csv"))


wta_long <- wta_df %>% pivot_longer(c(winner_name, loser_name),
                                    names_to = "won_or_lost",
                                    values_to = "player") %>%
  select(won_or_lost, player, w_ace, l_ace)

keep_n20 <- wta_long %>% group_by(player) %>% count() %>%
  filter(n > 20)

wta_tennis <- semi_join(wta_long, keep_n20, by = c("player"))

wta_tennis %>%
  select(won_or_lost, player, l_ace, w_ace) %>%
  group_by(player) %>%
  mutate(matches = n()) %>%
  ungroup() %>%
  group_by(player, w_ace, l_ace) %>%
  summarize(sum_w sum
```

Question 8:

part a: in this graph we are looking at averages. Plots that display relative length is not very good for plotting averages because we loose all information about variability

part b: 

```{r}
library(tibble)
set.seed(03092022)
toy_df <- tibble(group = c("a", "a", "a", "b", "b", "b", "c", "c",
                           "c", "c", "d", "d"),
       response = rnorm(12, 4, 3))

toy_mean <- toy_df %>%
  group_by(group) %>%
  summarise(mean = mean(response)) %>%
  mutate(group = fct_reorder(group, mean))

ggplot(data = toy_mean, aes(x = group,
                            y = mean)) +
  geom_point(colour = "blue", size = 3) +
  geom_point(data = toy_df, aes(x = group,
                                y = response)) +
  coord_flip() +
  labs(x = "Group",
       y = "Value")
```




